Compute Servers for Teaching Big Data

### 2025-02-04 paper.tex
.

    We compare and contrast different server options for providing undergraduate students with an authentic experience analyzing large, real world data sets located on powerful remote servers.
These server options include institutionally supported academic computers, commercial cloud services, public cloud services, and personal physical machines.
% TODO: Align these options with the sections
We briefly describe the course, and offer suggestions on when to choose which server option, based on the background of the instructor and the level of institutional support.
Looking to the future, we suggest requirements for a broadly shared educational compute service that will allow more instructors to teach similar classes without the overhead of having to manage and administer their own server.

@@

\subsection{Dedicated Support}

@@

The disadvantages of Jetstream2 for teaching college classes are the security model, the lack of safeguards, and the annual renewals.
Jetstream2 caters to the needs of individual researchers and small research groups, and they support this primary use case very well.
In contrast, teaching a semester long undergraduate class is certainly not Jetstream2's primary use case.
When you're awarded an allocation, every user gets complete access to every part of the account.
Students have administrative access to everything the instructor creates, and every other student's machine.
There are no limits on how much compute an individual can use, so it's possible for one user to use up all the credits.
In other words, the only thing that's stopping them from doing something bad is their own awareness and respect for others.
We mostly got around these issues by setting very clear expectations for behavior, with the understanding that this whole idea is based on trust.
Even so, a student unintentionally left a large machine running for several weeks.
We weren't monitoring as diligently as we should have, so this used up a significant portion of our 100,000 service units.
Fortunately, we still had plenty to get through the course.
Finally, our allocation on Jetstream2 requires annual renewal, which makes us hesitate to rely on it as a long term solution for teaching this class.
We expect they would renew every year, but there's no guarantee.

@@

If it was easier to set up and run a server, then we believe that more instructors would be interested in teaching this kind of class.
For a lone instructor with no support, the barriers to entry are significant.
We would welcome a solution that provides this capability publicly.
It must have the following features:
\begin{enumerate}
\item \textbf{command line access} If this is present, then it should be possible to install and use all manner of command line software, including R, Python, and SQL.
\item \textbf{100 GB data storage} 
\item \textbf{free or low cost} Free for students is ideal, of course.
    If that's not possible, then the cost over the semester should not exceed that of a reasonably priced textbook.
\end{enumerate}
Based on our observations above, there are several other features that would be nice to have, and we list some of them here.

\begin{enumerate}
\item \textbf{longevity} It's much easier for instructors to invest in a particular solution if they can be sure that it will be around in 5 years.
\end{enumerate}

>>>>

% 112 words in abstract
% 4500 words in document
    We compare and contrast different compute server options for teaching an undergraduate statistics class centered on analyzing large, real world data sets located on powerful remote servers.
These options include institutionally supported servers, personal physical servers, public cloud services, and commercial cloud services.
% TODO: Align these options with the sections
We briefly describe the course, share our experience with each of these options including advantages and disadvantages, mention some technical considerations, and suggest when to choose which option.
Looking to the future, we outline requirements for a broadly shared educational compute service that will allow more instructors to teach similar classes without the overhead of having to manage and administer their own server.

@@

\subsection{Institutional Support}

@@

The disadvantages of Jetstream2 for teaching college classes are the security model, the lack of safeguards, and the annual renewals.
Jetstream2 caters to the needs of individual researchers and small research groups, and they support this primary use case very well.
In contrast, teaching a semester long undergraduate class is certainly not Jetstream2's primary use case.
When you're awarded an allocation, every user gets complete access to every part of the account.
Students have administrative access to everything the instructor creates, and every other student's machine.
There are no limits on how much compute an individual can use, so it's possible for one user to use up all the credits.
In other words, the only thing that's stopping them from doing something bad is their own awareness and respect for others.
We mostly got around these issues by setting very clear expectations for behavior, with the understanding that this whole idea is based on trust.
Even so, a student unintentionally left a large machine running for several weeks.
We weren't monitoring as diligently as we should have, and this used up a significant portion of our 100,000 service units, but we still had plenty to get through the course.
Finally, our allocation on Jetstream2 requires annual renewal, which makes us hesitate to rely on it as a long term solution for teaching Stat 129.
We expect they would renew every year, but there's no guarantee.

@@

If it was easier to set up and run a server, then we believe that more instructors would be interested in teaching this kind of class.
For a lone instructor with a heavy teaching load and no support, the barriers to entry are significant.
We would welcome a solution that provides this capability publicly.
It must have the following features:
\begin{enumerate}
\item \textbf{command line access} If this is present, then it should be possible to install and use all manner of command line software, including R, Python, and SQL.
\item \textbf{100 GB shared data storage} It should be easy to share the same data with every student.
    1 TB would be better.
\item \textbf{free or low cost} Free for students is ideal, of course.
    If that's not possible, then the cost over the semester should not exceed that of a reasonably priced textbook.
\end{enumerate}
Based on our observations above, there are several other features that would be nice to have, and we list some of them here.

\begin{enumerate}
\item \textbf{longevity} It's much easier for instructors to invest in a particular solution if they can be sure that it will be around in 5 years.
\item \textbf{load balancing} Current cloud solutions don't do a great job.
\end{enumerate}
